\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-14 arch.tex
Updated execution plan figures :   - replaced \tau+1 by \tau' to avoid some confusion   - corrected the conflicting case as it was not a conflict previously Few typos and correction in  arch.tex

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation \comment{Might be good to articulate what ``one'' step
    looks like}. The returned value indicates if a complete plan has
  been found. If so this part \comment{what does 'this part' mean? Can
    we sketch a figure or make it more formal?} of the plan is
  provided to {\em external} timelines of this reactor.

Both these deliberative functions have different scope and temporal
constraints. The \texttt{synchronize} call is considered as atomic and
will focus only on state inference for the current tick by identifying
the current value of {\em internal} state variables from the latest
updates on the {\em external} state variables. This process propagates
up in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process which is
focused on the future evolution of state variables; it can take
multiple steps spanning multiple ticks for a reactor to produce a plan
that is both complete and valid. When such plan is identified
\comment{how? Is there a terminating condition?} its {\em external}
timeline(s) are then dispatched to the corresponding reactors which
own these times and which in turn can start their own deliberation
\texttt{step}s to satisfy these new objectives. Consequently execution
occurs as these goals and plans propagate down in the reactor
hierarchy until they eventually become goals for the lowest level
reactor which -- instead of deliberating -- may just transform this
goal into a command for execution in hardware.

@@

@@

The planning process inside a reactor is directly based on the way
europa work which was described in previous section. Still some simple
alterations were made in order to both include deliberation
information one reactor provide to the architecture -- such as its
deliberation latency and its planning look-ahead both expressed in
ticks -- and the fact that we potentially need to interleave this
deliberation with not the synchronization process but also with the
deliberation of other reactors\footnote{Current implementation of \rx
  is running on a single process and it is the responsibility of the
  agent itself to emulate the multi-threading of the reactors
  deliberation and synchronization}. For this reason while the overal
planning of a europa reactor strictly relies on the europa framework
plnanning solver, its slices the execution of Algorithm
\ref{alg:europa:solve} in atomic steps. The interruption is done
around the recursive call (line \ref{li:recurse}) while always
ensuring that the current partial plan mainitained is never proven
inconsistant at any single step. This can be managed by allowing the
call to backtrack until a DecisionPoInt that is not exhausted (ie
$decision.hasNext()$ is \texttt{true}) is find in the call stack.

@@

The most obvious case is how synchronization allow propagate of {\em
  external} observations as a new tick occur in the plan. In or
presentation of synchronization, we isolated the problem by stating
that we will assume at this point that the reactorhas no future goal
or more accurateely no plan to enforce at this stage. By doing so we
were able to develop the way we augment yhe europpa solver to be able
to propagete {\em external} observations in order to identify the
reactor {\em internal} state for the same tick. By doing so we were
able to focus on synchronization as a state evluation process that can
be resolved within the europa framework with few extensions. simliarly
we presented deliberation at this fairly high level not necessarily
discussing how the disruptive nature of synchronization in this
process could impact the way this planning process will evolve.

We first anlyse the case during which synchronization occurs after the
planning has vcompleted (ie the last seep of deliberation resulted
into a plan that is considered as complete for the current
look-ahead). As synchronization starts with the plan produced this
will gave a frame for the search process. Therefore this deliberation
will not only be a model-based estimation uniquely based on the {\em
  external} observations but will also include the reactor intentions
as described by the plan produced. Consider for example the Shopping
Agent model described during the europa presentation and that the
location of the \texttt{agent} is an {\em external} timeline. For
example as, the reactor add the goal to have \texttt{Own} milk and the
current {\em external} states indicated that it was \texttt{At(home)}
it produced the plan in Fig. \ref{fig:shop:exec0}. The {\em external}
part of the plan -- namely the tokens \texttt{Going(home,
  superMarket)} and \texttt{At(SuperMarket)} -- and integrated as
goals for the reactor managing this specific state varaiable {\em
  internally}.

@@

Now, consider that at the next tick $\tau+1$ the reactor managing the
agent location was able to change the location state varaibale to the
\texttt{Going}. \rx informs our reactor of this new observation which
is added in the plan structure as a fact -- starting exactly at
$\tau+1$ and lastign for an unlknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization the solver can identify the similarity between this
new observation oand the next token in the existing plan. Which mean
that instead of making a blind exhaustive search of all the possible
implication the synchronization process starts its execution with the
assumption that these new observations are consitent with the current
plan maintained. As a result at the end of synchronization we will
have the plan depicted in Fig. \ref{fig:shop:exec1}. Which just
propagated the information provided by the new observation in the
exisiting plan (restricting in turn the start time of \texttt{Going}
and by propagtion of the constraints \texttt{Go} to be {\em exactly}
$\tau+1$). While focusing uniquely on the execution frontier ($\tau+1$
in Fig. \ref{fig:shop:exec1}), the europa solver took benefit of the
exisiting plan to evaluate a solution that asumes that the observation
is consistent with what was decided inside this reactor.

This choice is directed not only by some natural heurisitc choices
made in the solver but even more enforced in general by the heuristic
order of the decision point reolution choices for our {\em internla}
state flaw presented in section \ref{sec:arch:sync} {\em\color{red}
  TODO: need to rework the presentation of thee choices so I can refer
  to them more directly -- algorithm or pseudo theorem should do the
  trick}. Should such a flaw be revealed during synchronization --
meaning that the current state of an {\em internal} state varaibale is
not fully grounded -- the sequencing of the choices will atempt first
a conservative choice in term of maintaining previous state, then an
optimisitic choice in term of the execution/advance in our current
plan to finally attempt other choices that were not predicted by the
model. The two first possible choices are strongly directed by the
exisiting plan which help direct the search within this frame. This
often allow to have a synchronization that is much more directed and
i, in nominal situation such as the one we showed resolve the
deliberation in few steps (often without any backtrack in the search).

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t1}
  \caption{Shopping example : Result of synchronization after the {\em
      Going} observation was received by the reactor with thew plan
    showed in Fig. \ref{fig:shop:exec0}.}
  \label{fig:shop:exec1}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax}  
  \caption{Illustration of a conflict suring synchronization and the
    resulting relaxed plan for recovery} 
  \label{fig:shop:relax}
\end{figure}

@@

By our design choices -- both at the architecture level and the
integration of the europa framework for embedded deliberation -- we
were able to implement an architecture that avoid this limitation
while remaining in a formal frame that still allow for deliberation
and model-based agent control. We believe that this tighter
integration of planning and execution allow the system to ba more
acute of its environment allowing better informaed decision and
tighter reaction. This is possible while still integrating information
such as a long latency for the reactors as long as one can ensure that
synchronization of al the reactors and deliberation steps can be done
in a time that is resonably small in comparison of the agent tick
duration.

>>>>

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of \eu for example pone step of
  deliberation correspod to resolve one sibngle flaw while ensuring
  that the resulting plan is not proven inconsistent. This implies
  that it also includes the backtracking of the search in such case. 
  The returned value by this step function is a boolean which
  indicates whether the reactor needs some extra step to deliberate or
  if this last step resulted on a complete plan for the reactor
  look-ahead. On the later, the part of the plan that concern this
  reactor {\em external} timelines is provided in order to be
  dispatched as goals to whichever reactor that is reponsible for
  these. \comment{By extending I may have saud things I developped
    below ... will have  to check}
\end{itemize}

Both these deliberative functions have different scope and temporal
constraints. The \texttt{synchronize} call is considered as atomic and
will focus only on state inference for the current tick by identifying
the current value of {\em internal} state variables from the latest
updates on the {\em external} state variables. This process propagates
up in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process which is
focused on the future evolution of state variables; it can take
multiple steps spanning multiple ticks for a reactor to produce a plan
that is both complete and valid. When such plan is identified
\comment{how? Is there a terminating condition? No more flaws but I
  washoping it would have been defined in \eu section} its {\em external}
timeline(s) are then dispatched to the corresponding reactors which
own these times and which in turn can start their own deliberation
\texttt{step}s to satisfy these new objectives. Consequently execution
occurs as these goals and plans propagate down in the reactor
hierarchy until they eventually become goals for the lowest level
reactor which -- instead of deliberating -- may just transform this
goal into a command for execution in hardware.

@@

For the sake of simplicity we introduce first how a reactor can track
and identify its state in the context where it does not have any
compelling need to deliberate. Assume here that we have a reactor that
have no future goal. Even in this context, it is a core requirement in
T-REX (see \cite{py10}) to have each reactor to produce at every
single tick their {\em internal} state or at least, for the ones not owning
{\em internal} timelines, make sure that {\em external} observations
are taken into account for every tick. This requirement is enforced to
ensure that all the reactors of an agent share a consistent view of
the world as time advance.

Taking into account {\em external} state updates is manage d by
introducing these observation in the plan structure as factsand ensure
that they remain compatible sewith the current plan mainainted by the
reactor. The more delicate part of synchronization is to decide of the 
{\em internla} state values for the current tick. In the Shopping
example for example we have the following code snippet whihch helps
identify when we can \texttt{Buy} a product:
\begin{verbatim}
Agent::Buy {
  eq(duration, 10);
  // [...]
  contained_by(condition object.location.At currLocation);
  // [...]
}
\end{verbatim}
If \texttt{object.location} is {\em external} and the \texttt{Agent}
is {\em internal}. I may encounter the situation where I am at the
location where I can buy one (possibly among others) product. Already
there's here  a decision problem related to the state value of the
\texttt{Agent} timeline where I can either {\em Buy} a specific
product now, wait a little longer before doing so or just simply not
buy any product at this location.

In the \eu based reactor we introduce this decision reuqirement 
as a new type of flaw for the \eu framework. This flaw enforces at the
reactor to identify fully its internal state for the current tick. By
using this new flaw we can describe the synchronization process as
this general sequence :

@@

The planning process inside a reactor is directly based on the way
\eu work which was described in previous section. Still some simple
alterations were made in order to both include deliberation
information one reactor provide to the architecture -- such as its
deliberation latency and its planning look-ahead both expressed in
ticks -- and the fact that we potentially need to interleave this
deliberation with not the synchronization process but also with the
deliberation of other reactors\footnote{Current implementation of \rx
  is running on a single process and it is the responsibility of the
  agent itself to emulate the multi-threading of the reactors
  deliberation and synchronization}. For this reason while the overal
planning of a \eu reactor strictly relies on the \eu framework
plnanning solver, its slices the execution of Algorithm
\ref{alg:europa:solve} in atomic steps. The interruption is done
around the recursive call (line \ref{li:recurse}) while always
ensuring that the current partial plan mainitained is never proven
inconsistant at any single step. This can be managed by allowing the
call to backtrack until a DecisionPoInt that is not exhausted (ie
$decision.hasNext()$ is \texttt{true}) is find in the call stack.

@@

The most obvious case is how synchronization allow propagate of {\em
  external} observations as a new tick occur in the plan. In or
presentation of synchronization, we isolated the problem by stating
that we will assume at this point that the reactorhas no future goal
or more accurateely no plan to enforce at this stage. By doing so we
were able to develop the way we augment yhe europpa solver to be able
to propagete {\em external} observations in order to identify the
reactor {\em internal} state for the same tick. By doing so we were
able to focus on synchronization as a state evluation process that can
be resolved within the \eu framework with few extensions. simliarly
we presented deliberation at this fairly high level not necessarily
discussing how the disruptive nature of synchronization in this
process could impact the way this planning process will evolve.

We first anlyse the case during which synchronization occurs after the
planning has vcompleted (ie the last seep of deliberation resulted
into a plan that is considered as complete for the current
look-ahead). As synchronization starts with the plan produced this
will gave a frame for the search process. Therefore this deliberation
will not only be a model-based estimation uniquely based on the {\em
  external} observations but will also include the reactor intentions
as described by the plan produced. Consider for example the Shopping
Agent model described during the \eu presentation and that the
location of the \texttt{agent} is an {\em external} timeline. For
example as, the reactor add the goal to have \texttt{Own} milk and the
current {\em external} states indicated that it was \texttt{At(home)}
it produced the plan in Fig. \ref{fig:shop:exec0}. The {\em external}
part of the plan -- namely the tokens \texttt{Going(home,
  superMarket)} and \texttt{At(SuperMarket)} -- and integrated as
goals for the reactor managing this specific state varaiable {\em
  internally}.

@@

Now, consider that at the next tick $\tau' = \tau+1$ the reactor managing the
agent location was able to change the location state varaibale to the
\texttt{Going}. \rx informs our reactor of this new observation which
is added in the plan structure as a fact -- starting exactly at
$\tau'$ and lastign for an unlknown duration which is at least 1
tick. As this new observation is integrated in the plan through
synchronization the solver can identify the similarity between this
new observation oand the next token in the existing plan. Which mean
that instead of making a blind exhaustive search of all the possible
implication the synchronization process starts its execution with the
assumption that these new observations are consitent with the current
plan maintained. As a result at the end of synchronization we will
have the plan depicted in Fig. \ref{fig:shop:exec1}. Which just
propagated the information provided by the new observation in the
exisiting plan (restricting in turn the start time of \texttt{Going}
and by propagation of the constraints \texttt{Go} to be {\em exactly}
$\tau'$). While focusing uniquely on the execution frontier ($\tau'$
in Fig. \ref{fig:shop:exec1}), the \eu solver took benefit of the
exisiting plan to evaluate a solution that asumes that the observation
is consistent with what was decided inside this reactor.

This choice is directed not only by some natural heurisitc choices
made in the solver but even more enforced in general by the heuristic
order of the decision point reolution choices for our {\em internla}
state flaw presented in section \ref{sec:arch:sync} {\em\color{red}
  TODO: need to rework the presentation of thee choices so I can refer
  to them more directly -- algorithm or pseudo theorem should do the
  trick}. Should such a flaw be revealed during synchronization --
meaning that the current state of an {\em internal} state varaibale is
not fully grounded -- the sequencing of the choices will atempt first
a conservative choice in term of maintaining previous state, then an
optimisitic choice in term of the execution/advance in our current
plan to finally attempt other choices that were not predicted by the
model. The two first possible choices are strongly directed by the
exisiting plan which help direct the search within this frame. This
often allow to have a synchronization that is much more directed and, 
in nominal cases such as the one we showed resolve the deliberation 
in few steps (often without any backtrack in the search).

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t1}
  \caption{Shopping example : Result of synchronization after the {\em
      Going} observation was received by the reactor with the plan
    showed in Fig. \ref{fig:shop:exec0}.}
  \label{fig:shop:exec1}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax}  
  \caption{Illustration of a conflict during synchronization and the
    resulting relaxed plan for recovery} 
  \label{fig:shop:relax}
\end{figure}

@@

By our design choices -- both at the architecture level and the
integration of the \eu framework for embedded deliberation -- we
were able to implement an architecture that avoid this limitation
while remaining in a formal frame that still allow for deliberation
and model-based agent control. We believe that this tighter
integration of planning and execution allow the system to ba more
acute of its environment allowing better informaed decision and
tighter reaction. This is possible while still integrating information
such as a long latency for the reactors as long as one can ensure that
synchronization of al the reactors and deliberation steps can be done
in a time that is resonably small in comparison of the agent tick
duration.
